### 2013-07-18 2013specconstr/Main.tex
specconstr paper

\documentclass{tmr}
\usepackage{style/utils}
\usepackage{style/code}

\title{Constructor specialisation in GHC}
\author{Amos Robinson\email{amosr@cse.unsw.edu.au}}

\begin{document}
\makeatactive

\begin{introduction}
In Haskell, loops, even those of higher-order combinators such as @map@, are expressed as recursive functions. 
With boxed objects such as @Int@, each iteration ends up allocating new objects in memory, only to be deconstructed immediately in the next iteration.
For high-performance code, these spurious allocations turn out to have serious penalties.

This problem has been solved specifically for boxed objects such as @Int@, @Float@ and others in the worker/wrapper transform~\CITE.
However, in the case of code produced by stream fusion~\CITE, other types such as @Either@ and tuples are used as loop state.
Constructor specialisation, or SpecConstr, is a generalisation of worker/wrapper for arbitrary types.
This is able to remove such allocations in stream-fusion code, and in some circumstances, produce object code that competes with hand-written C.
\end{introduction}

\section{}

SpecConstr

ForceSpecConstr, termination

And not specialising ``too many times'' on recursive types

Code blowup

Seeding specialisation of non-exported functions

\end{document}