### 2015-02-02 2015mergingmerges/section/02-Streams.tex
ok

%!TEX root = ../Main.tex
\section{Streams and combinators}
\label{s:Streams}
First, we define an abstraction of two-input combinators, very similar to stream fusion\CITE or co-iterative stream functions of Caspi and Pouzet\cite{caspi1998co}.
We present a simplified stateless version here, which would not be able to express all useful combinators (such as segmented scan) but suffices for merge, append and zip.
We also only treat combinators that consume both input streams linearly, which excludes combinators like cross product or gather.
In Coq\footnote{The definitions have been reordered for clarity}:

\begin{code}
Definition Stream2 (a b c : Set)
 := option a -> option b -> Step2 c

Inductive Step2 c :=
 | MkStep2 : Move -> Move -> option c -> Step2 c
 | Done2   : Step2 c.

Inductive Move :=
 | Forward
 | Stay.
\end{code}

At each evaluation step, a @Stream2@ is called with the value at each input stream, or @None@ if the input stream is finished.
The return of each @Stream2@ evaluation step can be either a @MkStep2@ indicating that further processing is possible, or @Done2@, in which case the combinator has finished.
A @MkStep2@ contains two @Move@s, which designate which input streams should be consumed and the next value pulled, and which should stay where they are.
@MkStep2@ also contains an optional output value.

The main difference from stream fusion is that the coordination between streams is lifted out.
In stream fusion, each stream transformer takes a stream as input and evaluates it itself.
This means that, since the stream evaluation is local to a stream, scheduling decisions can only be made locally.
These local schedules cannot guarantee desired properties like bounded memory, as multiple streams may need to be stepped together if they use the same input.
Instead, the co-iterative version is evaluated by an outside scheduler, which can find an evaluation order that satisfies these properties.

For single-input combinators such as map, filter and scan, @Stream1@ is defined below.
\begin{code}
Definition Stream1 (a b : Set)
 := option a -> Step2 b

Inductive Step1 b :=
 | MkStep1 : Move -> option b -> Step1 b
 | Done1   : Step1 b.
\end{code}

\subsection{Examples}

Below, we show example combinators.
The first combinator pairwise zips two streams together, consuming both streams in lock-step.

\begin{code}
Definition Zip (A B : Set)
 :  Stream2 A B (A*B)
 := fun a b =>
    match (a,b) with
    | (Some a, Some b)
    => MkStep2 Forward Forward (Some (a,b))
    | _               
    => Done2
    end.
\end{code}

Next, append consumes from the left stream until it is empty, then consumes the right stream.
\begin{code}
Definition Append (A : Set)
 :  Stream2 A A A
 := fun a b =>
    match (a,b) with
    | (Some a, _)
    => MkStep2 Forward Stay    (Some a)
    | (None,   Some b)
    => MkStep2 Stay    Forward (Some b)
    | (None,   None)               
    => Done2
    end.
\end{code}

As a final example, the merge join inspects both its input elements, and consumes the one with the smallest value.
There are many variations on this theme, but this particular version that outputs all values in order is actually equivalent to a segmented append.

\begin{code}
Definition MergeJoin (A : Set)
 :  Stream2 (nat*A) (nat*A) (nat*A)
 := fun a b =>
    match (a,b) with
    | (Some (ia,a), Some (ib, b))
    => if   le_dec ia ib
       then MkStep2 Forward Stay (Some (ia, a))
       else MkStep2 Stay Forward (Some (ib, b))
    | (Some (ia,a), None)
    =>      MkStep2 Forward Stay (Some (ia,a))
    | (None, Some (ib,b))
    =>      MkStep2 Stay Forward (Some (ib,b))
    | (None, None)
    =>      Done2
    end.
\end{code}

\subsection{Kahn process networks}

We define a simplified Kahn process network that contains no cycles (recursion).

\begin{code}
Inductive Graph
 : list Set -> list Set -> list Set -> Type :=
\end{code}
The graph is parameterised by the list of inputs, the list of interior nodes, and the list of outputs.

\begin{code}
 | Empty  : Graph [] [] []

 | Inp    : forall a i n o,
            Graph i n o
         -> Graph (a::i) (a::n) o

 | Out    : forall a i n o,
            Index n a
         -> Graph i n o
         -> Graph i n (a::o)
\end{code}

A @Graph@ can be @Empty@, in which case there are no inputs, nodes or outputs.
The @Inp@ constructor takes an existing graph, and adds an input node to the start.
Because the graph cannot be recursive, the inner graph of @Inp@ cannot refer to the newly created @Inp@.
Similarly, @Out@ takes an existing graph, and an index into the list of interior nodes, and marks that node as an output.

\begin{code}
 | Trans1 : forall a b i n o,
            Index n a
         -> Stream1 a b
         -> Graph i n o
         -> Graph i (b::n) o.

 | Trans2 : forall a b c i n o,
            Index n a
         -> Index n b
         -> Stream2 a b c
         -> Graph i n o
         -> Graph i (c::n) o.
\end{code}
The two transition constructors, @Trans1@ and @Trans2@ take @Stream1@ and @Stream2@ respectively, which defines the actual computation of the transition. They also extend existing graphs, and their inputs must be in the existing graph, again disallowing recursive edges or cycles.