### 2015-03-02 2015mergingmerges/Main.tex
ok

\documentclass{sigplanconf}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{mathptmx}
\usepackage{stmaryrd}
\usepackage{hyperref}
\usepackage{alltt}
\usepackage{url}
\usepackage{float}
\usepackage{style/utils}
\usepackage{style/code}
\usepackage{style/proof}
\usepackage{style/judgements}

% -----------------------------------------------------------------------------
\begin{document}

% \exclusivelicense
% \conferenceinfo{}{}
% \copyrightyear{2015}
% \copyrightdata{}
\doi{}
% \pagenumbering{gobble}

\title{Merging merges, more or less}

\authorinfo{
  Amos Robinson$^\dagger$ 
  \and Ben Lippmeier$^\dagger$
  \and Gabriele Keller$^\dagger$ 
  \and Manuel Chakravarty$^\dagger$ 
}{
  \vspace{5pt}
  \shortstack{
    $^\dagger$Computer Science and Engineering \\
    University of New South Wales, Australia \\[2pt]
    \textsf{\{amosr,benl,keller,chak\}@cse.unsw.edu.au}
  }
}

\maketitle
\makeatactive

\begin{abstract}
Fusion is an essential part of optimising high-level array computations.
Fusion systems for combinators tend to focus on combinators with data-independent access patterns such as @zip@, @filter@ and @map@, while ignoring @merge@ as in merge sort.
In this paper, we show that describing combinators as deterministic finite automata allows us to express more combinators, while leading to a fusion algorithm based on intersection of automata.
\end{abstract}

\category
	{D.3.4}
	{Programming Languages}
	{Processors---Compilers; Optimization}

\terms
	Languages, Performance

\keywords
	Arrays; Fusion; Haskell

\input{section/01-Introduction.tex}
\input{section/02-Streams.tex}
\input{section/03-Access.tex}
\input{section/XX-RelatedWork.tex}

\section*{Acknowledgements}
The audience of the inaugural SAWDAP workshop, especially Thomas Sutton and Patryk Zadarnowski, were very helpful in an early version of this work.

\bibliographystyle{plain}
\bibliography{Main}

\end{document}