### 2015-11-14 2016icicle/section/01-Introduction.tex

%!TEX root = ../Main.tex
\section{Introduction}
\label{s:Introduction}

For queries over a large amount of data, it is important that the queries are as fast as possible.
If a query must read and loop over the data multiple times, that could more than double the query time, as well as potentially requiring large intermediate storage.
It also turns out that with some ingenuity, many interesting algorithms can be expressed or at least approximated in a single pass.
The problem with many existing query languages, however, is that it can be hard to tell how many passes will be used for a given query.
This execution-agnostic property of declarative languages can be both a benefit and a disadvantage.

With a language like SQL, it may not be immediately obvious how many passes a query requires, and it may depend on particular compiler optimisations and the mood of the database.
For example, how many iterations do the following queries require?

\begin{code}
SELECT open - MIN(open)    FROM stocks;
SELECT open - MIN(open, 1) FROM stocks;
SELECT        MIN(open)    FROM stocks;
SELECT open,  MIN(open)    FROM stocks;
\end{code}

For our purposes, we would rather stricter compile-time guarantees: if a query typechecks and compiles, it should be able to be executed efficiently.
The aim is to remove the mystery of database and compiler optimisations with a clear, unambiguous definition of which queries should be accepted and efficient, and which will be rejected.

\TODO{Look into data flow languages like Lucid and how they enforce causality, cite them}

We make the following contributions:
\begin{itemize}
\item 
We present a query language, Icicle, with a type system using modality types\CITE{Rowan} to enforce the single-pass restriction(\S\ref{s:Source}).
\item
We present an intermediate language similar to MapReduce\CITE{?} that simplifies fusion between multiple queries(\S ref).
\item
We describe an algorithm for extracting efficient imperative code from the intermediate language, used to extract C code in our implementation (\S ref).
\item
We show benchmarks of our implementation, which performs favourably even compared to hand-optimised text programs such as @grep@ and @wc@ (\S ref).
\end{itemize}

Our implementation is available at \url{https://github.com/ambiata/icicle}.