### 2016-03-01 2016icicle/section/04-Core.tex

%!TEX root = ../Main.tex
\section{Core language}
\label{s:Core}

This section presents the Core language, which we use as an intermediate language before converting to imperative code.

\subsection{Grammar}

\input{figures/Core-Grammar.tex}

% \input{figures/Core-Compile.tex}