### 2016-04-06 2016fhpc/section/02-Core.tex
ok

%!TEX root = ../Main.tex
\section{Icicle Core}
\label{s:IcicleCore}

Horizontal fusion for imperative loops is relatively easily, when the loops have the same number of iterations.
The problem with imperative loops occurs when trying to remove duplicate computations.
Firstly, the `definition' of a computation is split across multiple places: because the initial value of an accumulator and its update expression are in separate parts of the program, some analysis must be done to recover these.
For example, in the following loop, @a@ and @b@ have equivalent `update' expressions, but could denote very different computations depending on the implementation of @function@.
\begin{code}
a = 0;
b = 1;
for (...) {
  a = function(a);
  b = function(b);
}
\end{code}

Secondly, the order of statements in a loop affects the meaning: when accumulators are mutually dependent on each other, one can reorder the statements to produce a different, but still valid program.
These two loops are not equivalent:
\begin{code}
a = 0;
b = 1;
for (...) {
  a = a + b;
  b = a - b;
}
for (...) {
  b = a - b;
  a = a + b;
}
\end{code}

Neither of these problems are insurmountable, but the analysis required certainly complicates our goal of removing duplicate computation after fusion.

With this in mind, we will introduce our intermediate language based on what we call ``fold nests'', as opposed to loop nests.
Firstly, each accumulator should be defined in one place: this means that if two accumulators are alpha equivalent, they are equivalent.
Secondly, reordering statements (or accumulators) should not affect the meaning: two programs are equivalent if they contain equivalent sets of bindings, regardless of ordering.

We now convert the first example to folds:
\begin{code}
a = 0;
b = 1;
for (...) {
  a = function(a);
  b = function(b);
}

==>
loop ... {
  stage {
    fold a = 0
        then function(a);
  }
  stage {
    fold b = 1
        then function(b);
  }
}
\end{code}

Each fold binding must be grouped inside a @stage@ block, which allows mutual recursion across folds.
In this case, there is no mutual recursion, so the stages only contain one binding each.
The first fold is given the name @a@ with @fold a@.
Its initial value comes straight after the equals sign, so @a@ starts with a value of @0@.
The update or `kons' expression comes after @then@, and in the update expression any reference to @a@ refers to the current value of the fold.

It now becomes easier to see that the two folds are not equivalent, as the entire definitions are in one place.

The next example requires mutual recursion.
\begin{code}
a = 0;
b = 1;
for (...) {
  a = a + b;
  b = a - b;
}

==>
loop ... {
  stage {
    fold a = 0
        then a + b;
    fold b = 1
        then a' - b;
  }
}
\end{code}

Here, the update expression for @a@ refers to a later fold, @b@.
Update expressions can refer to later fold bindings in the same stage, as well as all earlier fold bindings.
The value referred to is the value of the binding at the start of the iteration, before any updates have occurred.
The new value of earlier folds, after being updated, can be referred to by suffixing a prime to the name: in the binding for @b@, the updated @a@ is referred to by @a'@.
This restriction of only accessing the new value of earlier folds is akin to the causality restriction in dataflow languages\cite{mandel2010lucy}.

Another way to think of the new value restriction above is disallowing cycles in the references to new values:
when @b@ references @a'@, the new value of @a@, it means that @a'@ must be computed before @b'@ can be.
If there were a cycle where @a'@ also depended upon @b'@, there would be no order to compute them in.
Disallowing cycles ensures an execution order.

By making the distinction between `old value' and `new value' explicit, the ordering in the program no longer has any effect on the semantics.
After converting the reordered loop with different semantics, the two programs are no longer alpha-equivalent:
\begin{code}
a = 0;
b = 1;
for (...) {
  b = a - b;
  a = a + b;
}

==>
loop ... {
  stage {
    fold b = 1
        then a - b;
    fold a = 0
        then a + b';
  }
}
\end{code}

\begin{figure}

\begin{code}
N     := a,b,...
FoldN := a,b,...

Exp   := N
       | FoldN  | FoldN'
       | Exp Exp
       | \N. Exp

Bind  := fold FoldN : T = Exp then Exp;
       | let  N     : T = Exp;

Stage := stage { Bind...  }
Query := loop N : T { Stage... }

\end{code}

\caption{Grammar}
\label{fig:grammar}
\end{figure}

\newcommand\TypecheckX[3]
{                   #1
        ~\vdash~    #2~:~#3
}

\newcommand\TypecheckS[3]
{                   #1
        ~\vdash~    #2~\dashv~#3
}

\begin{figure}

$$
\boxed{\TypecheckS{\Gamma}{\mi{Query}}{\Gamma'}}
$$

$$
\ruleAx
{
  \TypecheckS{\Gamma}{@loop@~x~:~\tau~\{\}}{\emptyset}
}{QueryNil}
$$

$$
\ruleIN
{
  \TypecheckS{\Gamma,~x~:~\tau}{s}{\Gamma'}
  \quad
  \TypecheckS{\Gamma,~\Gamma'}{@loop@~x~:~\tau~\{s'\}}{\Gamma''}
}{
  \TypecheckS{\Gamma}{@loop@~x~:~\tau~\{s;~s'\}}{\Gamma' \cup \Gamma''}
}{QueryCons}
$$

$$
\boxed{\TypecheckS{\Gamma}{\mi{Stage}}{\Gamma'}}
$$

$$
\ruleAx
{
  \TypecheckS{\Gamma}{@stage@~\{\}}{\emptyset}
}{StageNil}
$$

$$
\ruleIN
{
  \TypecheckS{\Gamma,~x~:~\tau,~x'~:~\tau}{@stage@~\{s'\}}{\Gamma'}
  \quad
  \TypecheckX{\Gamma}{z}{\tau}
  \quad
  \TypecheckX{\Gamma~\cup~\{v~|~v'~\in~\Gamma'\}}{k}{\tau}
}
{
  \TypecheckS{\Gamma}{@stage@~\{@fold@~x~:~\tau~=~z~@then@~k;~s'\}}{\Gamma',~x~:~\tau,~x'~:~\tau}
}{StageFold}
$$

$$
\ruleIN
{
  \TypecheckS{\Gamma,~x~:~\tau}{@stage@~\{s'\}}{\Gamma'}
  \quad
  \TypecheckX{\Gamma}{x}{\tau}
}
{
  \TypecheckS{\Gamma}{@stage@~\{@let@~x~:~\tau~=~e;~s'\}}{\Gamma',~x~:~\tau}
}{StageLet}
$$

\caption{Typing rules}
\label{fig:typing}
\end{figure}

The full grammar is in figure~\ref{fig:grammar}, however primitives and types have been omitted.
It is assumed that there are some primitives such as @1@, @+@ and types such as @Int@, but the elements do not actually matter.

The typing rules are in figure~\ref{fig:typing}.
The rules shown only cover the staging and folds.
As with the grammar, we assume there is a typing rule $\TypecheckX{\Gamma}{\mi{Exp}}{\tau}$ which checks an expression under the given environment.