### 2016-05-12 2016fhpc/Main.tex
introduction and first section (Source), start on Core section

\documentclass[preprint]{sigplanconf}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{mathptmx}
\usepackage{mathtools}
\usepackage{stmaryrd}
\usepackage{hyperref}
\usepackage{alltt}
\usepackage{url}
\usepackage{float}
\usepackage{style/code}
\usepackage{style/proof}
\usepackage{style/utils}
\usepackage{style/judgements}

% -----------------------------------------------------------------------------
\begin{document}

% \exclusivelicense
% \conferenceinfo{}{}
% \copyrightyear{2015}
% \copyrightdata{}
\doi{}
% \pagenumbering{gobble}

\title{Icicle: write once, run once}

\authorinfo{
  Amos Robinson$^\dagger$$^\ddagger$
  \and Ben Lippmeier$^\dagger$
}{
  \vspace{5pt}
  \shortstack{
    $^\dagger$Computer Science and Engineering \\
    University of New South Wales, Australia \\[2pt]
    \textsf{amosr,benl@cse.unsw.edu.au}
  }
  \shortstack{
    $^\ddagger$Ambiata      \\
    Grow with data          \\[2pt]
    \textsf{amos.robinson@ambiata.com}
  }
}

\maketitle
\makeatactive

\begin{abstract}
When querying a large amount of data, simply iterating over the data may take hours.
If multiple queries are to be performed, it is important that work is not duplicated.
Queries that can be performed together must be performed in the same iteration.
We introduce a simple streaming language for computing queries in a single-pass over the data.
With type and other restrictions we guarantee fusion between queries on the same input table, before extracting efficient C code.
\end{abstract}

\category
	{D.3.4}
	{Programming Languages}
	{Processors---Compilers; Optimization}

\terms
	Languages, Performance

\keywords
	Arrays; Fusion

\input{section/01-Introduction.tex}
\input{section/02-Source.tex}
\input{section/03-Core.tex}
\input{section/04-Benchmarks.tex}
\input{section/05-Conclusion.tex}
\input{section/XX-Elimination.tex}
\input{section/XX-Related.tex}

\section*{Acknowledgements}

\bibliographystyle{plain}
\bibliography{Main}

\end{document}