### 2016-05-31 2016fhpc/Main.tex
minor changes

\documentclass[preprint]{sigplanconf}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{mathptmx}
\usepackage{mathtools}
\usepackage{stmaryrd}
\usepackage{hyperref}
\usepackage{alltt}
\usepackage{url}
\usepackage{float}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{style/code}
\usepackage{style/proof}
\usepackage{style/utils}
\usepackage{style/judgements}

% -----------------------------------------------------------------------------
\begin{document}

% \exclusivelicense
% \conferenceinfo{}{}
% \copyrightyear{2015}
% \copyrightdata{}
\doi{}
% \pagenumbering{gobble}

\title{Icicle: write once, run once}

\authorinfo{
  Amos Robinson$^{\alpha \beta}$
  \and Ben Lippmeier$^{\beta \gamma}$
}{
  \vspace{5pt}
  \shortstack{
    $^\alpha$Ambiata      \\
    \textsf{amos.robinson@ambiata.com}
  }
  \shortstack{
    $^\beta$UNSW Australia\\
    ~~~~\textsf{amosr,benl@cse.unsw.edu.au}
  }
  \shortstack{
    $^\gamma$Vertigo Technology \\
    ~~\textsf{benl@vergo.co}
  }
}

\maketitle
\makeatactive

\begin{abstract}
We present Icicle, a pure streaming query language which statically guarantees that multiple queries over the same input stream will be fused. We use a modal type system to ensure that fused queries can be computed in an incremental fashion, and a fold-based intermediate language to compile down to efficient C code. We present production benchmarks demonstrating significant speedup over existing queries written in R, as well as the widely used Unix tools \texttt{grep} and \texttt{wc}.

% When querying a large amount of data, simply iterating over the data may take hours. If multiple queries are to be performed, it is important that work is not duplicated. Queries that can be performed together must be performed in the same iteration. We introduce a simple streaming language for computing queries in a single-pass over the data. With type and other restrictions we guarantee fusion between queries on the same input table, before extracting efficient C code.
\end{abstract}

\input{section/01-Introduction.tex}
\input{section/02-Source.tex}
\input{section/03-Core.tex}
\input{section/04-Benchmarks.tex}
% \input{section/05-Conclusion.tex}
% \input{section/XX-Elimination.tex}
\input{section/XX-Related.tex}

\section*{Acknowledgements}

\bibliographystyle{plain}
\bibliography{Main}

% \input{section/A1-ExampleDerivation.tex}

\end{document}