### 2017-01-13 2017mergingmerges/section/XX-Related.tex
prose

%!TEX root = ../Main.tex
\section{Related work}

Existing systems for combinator fusion only support a subset of combinators, and only in limited cases.
Pull fusion supports joins where one combinator pulls from multiple inputs, but not splits where one output is used my multiple combinators.
Pull can thus express @merge@, but not @partition@.
Push fusion supports splits (@partition@), but not joins (@merge@).
Polarised fusion~\cite{lippmeier2016polarized} addresses these shortcomings by separating the computation into pull producers and push consumers, with explicit drain loops to convert between the two sections. However this requires manual plumbing.

Another way to characterise fusion systems is whether they are \emph{local} (shortcut) or \emph{global} fusion.
Local fusion systems such as stream fusion~\cite{coutts2007stream} use local rewrite rules to remove intermediate arrays, and tend to rely on general purpose compiler optimisations to expose opportunities for these rewrites to occur.
Generally, the local rewrite rule can only be applied once the producer has been inlined into the consumer.
Because of the local nature of these rewrites and the reliance on inlining, local fusion can only perform producer-consumer fusion with a single consumer, as inlining the producer into multiple consumers would duplicate work.
Local fusion is fragile because heuristics are used to determine whether inlining occurs~\cite{lippmeier2013data}.
This fragility poses serious problems to programmers who require fusion for adequate performance, as one must have deep knowledge of the internal compiler phases and library design in order to predict whether fusion will occur -- or worse yet, they must scour the intermediate code to count the number of loops produced. 
In contrast, global fusion systems such as those traditionally used in imperative compilers make use of specific optimisations which must be implemented separately, but are able to perform horizontal fusion and are less prone to the fragility of local fusion.

This local/global distinction is not a hard classification, as some systems fit somewhere in between.
For example, {\bf strymonas}~\cite{kiselyov2016stream} uses staged compilation for introspection of the code and to ensure appropriate inlining, but only performs producer-consumer fusion with a single local rule.
While not explicitly mentioned in the paper, Kiselyov et al~\cite{kiselyov2016stream}'s {\bf strymonas} system only allows streams to be used once.
Furthermore, the entire stream computation must be terminated with a single fold which takes a single stream.
This means multiple outputs with different rates cannot be treated, for example @partition@ which returns multiple arrays filtered with different predicates.

This innocent-looking yet useless program, which should pair each line with itself, actually pairs each line with the successive line.
\begin{code}
let lines = linesFromFile filename
in  zip lines lines
\end{code}

One would expect the above program to have the same behaviour as the following - or at least produce an error otherwise:
\begin{code}
zip (linesFromFile filename) (linesFromFile filename)
\end{code}

The one feature they support that we do not is concatMap/flatMap, which concatenates all produced substreams.
This is a limitation in the \emph{conversion} to process calculus, not in the process calculus itself.
Our process calculus can handle concrete instantiations of concatMap when specialised to a particular subprocess, but not the general case where the subprocess is statically unknown.
By using a staging restriction as they do, we could modify the conversion to create a specialised version of concatMap for the given argument.

\subsection{The problem with synchronised product}
Related work on Network Fusion~\cite{fradet2004network} allows fusion of Kahn Process Networks using synchronised product to fuse pairs of processes together.
Synchronised product is: both machines can take independent steps so long as it is not in the alphabet of the other, but have to agree on any shared actions.
So if both machines pull from @xs@, the @pull xs@ message can only be executed when both machines agree.

This sort of coordination is a bit too coarse and causes deadlocks.
This example cannot be fused by synchronised product: one zip is trying to pull from @as@ and the other from @bs@, but neither can proceed without the other.
Note that while this contrived example is unlikely to be written by hand, it is plausible that it would be part of a larger program after inlining has occurred.

\begin{code}
zips :: [a] -> [b] -> [a*b] * [b*a]
zips as bs =
  let abs = zip as bs
      bas = zip bs as
  in  abs, bas
\end{code}

When stream programs are encoded as labelled transition systems, some fusion can be done by computing the synchronised product of two transition systems, but this suffers deadlock when the two programs share multiple inputs and read them in different orders.
By annotating transition states with the status of each input channel, our fusion algorithm works for these cases where synchronised product does not.

% After inlining the definition of zip, we have:
% \begin{code}
% zips :: [a] -> [b] -> [a*b] * [b*a]
% zips as bs =
%   let abs = process abs.
%        let loop1 =
%           a <- pull as
%           b <- pull bs
%           push abs (a,b)
%           drop as
%           drop bs
%           jump loop1
%        in jump loop1
%       bas = process bas.
%        let loop2 =
%           b <- pull bs
%           a <- pull as
%           push abs (b,a)
%           drop bs
%           drop as
%           jump loop2
%        in jump loop2
%   in  abs, bas
% \end{code}
% 
% Here, in order to compute the synchronised product of these two, we need to know which are the common or shared actions.
% Shared are: @pull as@, @pull bs@, @drop as@ and @drop bs@.
% So both processes have to execute these together: if one wants to @pull as@, it can only act when the other one wants to @pull as@ as well.
% Thus, the two machines are deadlocked, waiting for the other one to agree.
% 
% For this case, a simple solution suffices to fix this: adding extra processes for splits. So we would add a new combinator @dup@ which takes one input stream and creates two output streams. It reads a single element from the input and pushes it to both outputs.
% 
% \begin{code}
% dup = stream_1_2 \is ols ors.
%   letrec
%     p1   = pull is p2
%     p2 i = push ols i (p3 i)
%     p3 i = push ors i p4
%     p4   = drop is p1
%   in p1
% \end{code}
% And the modified zip uses this like so:
% 
% \begin{code}
% zips :: [a] -> [b] -> [a*b] * [b*a]
% zips as bs =
%   let as1, as2 = dup as
%       bs1, bs2 = dup bs
%       abs = zip as1 bs1
%       bas = zip bs2 as2
%   in  abs, bas
% \end{code}
% 
% This has removed the deadlock, allowing a trace such as:
% \begin{code}
% as, as1, bs, bs1, bs2, as2
% \end{code}
% 
% However it is not hard to modify the program so that the deadlock still exists.
% By changing the each zip to pull from @as2@ and @bs2@ first, the streams are still used linearly, but deadlock occurs as @as1@ cannot be produced until after @bs2@ has been consumed, and vice versa.
% \begin{code}
% zips :: [a] -> [b] -> [a*b] * [b*a]
% zips as bs =
%   let as1, as2 = dup as
%       bs1, bs2 = dup bs
%       abs = zip as2 bs1
%       bas = zip bs2 as1
%   in  abs, bas
% \end{code}
% 
% Using @dup@ we can concoct a simpler example with two processes and one input stream where synchronised product deadlocks.
% \begin{code}
% zipself :: [a] -> [a*a]
% zipself as =
%   let as1, as2 = dup as
%       aas = zip as2 as1
%   in  aas
% \end{code}
% Here @dup@ is attempting to push to @as1@, but cannot until @zip@ pulls, while @zip@ is attempting to pull from @as2@, which cannot proceed until the @dup@ pushes.