### 2017-01-14 2017mergingmerges/section/XX-FutureWork.tex
prose explanation of finite streams extension

%!TEX root = ../Main.tex
\section{Extensions and future work}
\label{s:FutureWork}

We now discuss some of the shortcomings of the system, and extensions and future work to ameliorate this.

\subsection{Finite streams}
\label{s:Finite}

The processes we have seen so far deal with infinite streams, but in practice most streams are finite.
Certain combinators such as @fold@ and @append@ only make sense on finite streams, and others like @take@ produce inherently finite output.
We have focussed on the infinite stream version because it is somewhat simpler to explain and prove, but the extensions required to support finite streams do not require substantial conceptual changes.

We now describe the extensions required to support finite streams.
We add a new @closed@ constructor to the \InputState~ to encode the end of the stream.
Once an input stream is in the closed state, it can never change to another state: it remains closed thereafter.

We modify the @pull@ instruction so that it has two output labels (like @case@).
The first label, the read branch, is executed as before when the pull succeeds and a value is read from the stream.
The second label, the close branch, is executed when the stream is closed, and no more values will ever be available.
After a pull takes the close branch, any subsequent pulls from that stream will also take the close branch.

We add two new instructions for closing output streams and disconnecting from input streams.
Closing an output stream $(@close@~\Chan~\Goto)$ is similar to pushing an end-of-file marker to all readers.
As with @push@, the evaluation semantics of @close@ can only proceed if all readers are in a position to accept the end-of-file, but instead of setting the new \InputState~ to @pending@ with a value, the \InputState~ is set to @closed@.
After a stream has been closed, no further values can be pushed.

Disconnecting from input streams $(@disconnect@~\Chan~\Goto)$ signals that a process is no longer interested in the values of a stream.
This can be used when a process requires the first values of a stream, but does not require the whole stream.
If a process read the first values of a stream and then stopped pulling, its \InputState~ buffer would fill up and never be cleared, so no other process would be able to continue pulling from that stream.
Disconnecting the stream allows other processes to use the stream without the disconnected process getting in the way of computation.
The evaluation semantics for @disconnect@ remove the channel from the inputs of the process.
After removing the channel from the inputs, when a writing process tries to inject values, this process will just be ignored rather than inserting into the \InputState~ buffer and potentially causing writing to block.
After a process disconnects from an input channel, it can no longer pull from that channel.

We also add an instruction for terminating the process (@done@).
After all input streams have been read to completion or disconnected and output streams closed, the process may execute @done@ to signal that processing is complete.

The fusion definition must be extended to deal with these new instructions.
The static input state has a @closed@ constructor added and disconnection is encoded by removal from the input state, and the \ti{tryStep} changes more or less follow the evaluation changes.
Shared and connected pulls now deal with two more possibilities in the input state: the input may be closed in which case the close branch of the pull is taken; or the other process may have disconnected in which case the pull is executed as in the non-shared non-connected case.
Connected pushes must also deal with when the other process has disconnected in which case the push is executed as if it were non-connected.
For @in1@ and @out1@ channels, the new @close@ and @disconnect@ instructions are used as normal with no coordination required.
For connected @close@, as with @push@, the receiving process must have @none@ and the next step performs the @close@ and sets the input state to @closed@.
For shared @disconnect@, the @disconnect@ is only performed after both processes have disconnected; otherwise the entry is just removed from the input state.
For connected @disconnect@, the @disconnect@ is not performed and the entry is removed from the input state.

Finally, \ti{tryStepPair} is modified so that @done@ is performed when both machines are @done@.

These modifications allow our system to fuse finite streams as well as infinite.
We have implemented an initial prototype that supports finite streams, but future work is required to prove them correct.

\subsection{Fully abstract case interpretation}

This cannot be fused because it requires an unbounded buffer.
\begin{code}
zipltgts :: [a] -> [a*a]
zipltgts as =
  let as1 = filter (<0) as
      as2 = filter (>0) as
      aas = zip as1 as2
  in  aas
\end{code}

You might think the following can be fused.
It cannot because we treat @case@ conditions as fully abstract and make no attempt to filter out impossible combinations.
So while it cannot be true that the first filter reaches @>0 = true@ case and the second filter reaches @>0 = false@, we try both combinations and treat it as unfusable.
\begin{code}
zipgts :: [a] -> [a*a]
zipgts as =
  let as1 = filter (>0) as
      as2 = filter (>0) as
      aas = zip as1 as2
  in  aas
\end{code}