### 2017-02-25 2017mergingmerges/section/XX-Related.tex
related work

%!TEX root = ../Main.tex
\section{Related work}

This work aims to address the problems with existing combinator-based array fusion systems.
As stated in the introduction, neither pull-based or push-based fusion alone is sufficient.
Some combinators are inherently push-based, particularly those with multiple outputs such as @unzip@; while others are inherently pull-based, such as @zip@.

Short cut fusion is an attractive idea, as it allows fusion systems to be specified with a single rewrite rule.
However, short cut fusion relies on inlining which, like pull-based streams, only occurs when there is a single consumer.
Thus, short cut fusion is inherently biased towards pull fusion.
Push-based short cut fusion systems \emph{do} exist \cite{gill1993short}, but support only the intersection of pull and push: that is, neither @zip@ nor @unzip@ \cite{svenningsson2002shortcut,lippmeier2013data}.

Recent work on stream fusion by \citet{kiselyov2016stream} uses staged computation in a push-based system to ensure all combinators are inlined, but when streams are used multiple times this causes excessive inlining, which duplicates work.
For effectful input streams, for example reading from the network, this duplicated work can change the overall meaning of the program.
% I could write more about this eg only supporting a single output, but the other points probably apply to push streams in general

Data flow fusion~\cite{lippmeier2013data} is neither pull-based nor push-based, and supports arbitrary splits and joins.
It supports standard combinators such as @map@, @filter@ and @fold@, and converts each stream to a series with explicit rate types, similar to the clock types of Lucid Synchrone \cite{benveniste2003synchronous}.
These rate types ensure that well-typed programs can be fused without introducing unbounded buffers.
This allows unfusable programs to be caught at compile time.
The main drawback of data flow fusion is that it only supports a limited set of combinators, and adding more is non-trivial.

One way to address the difference between pull and push streams is to explicitly support both separately, as seen in \citet{bernardy2015duality} and \citet{lippmeier2016polarized}.
Here, pull streams have the type @Source@ and are represented as a source that is always available to be pulled from, while push streams have the type @Sink@ and are represented as a sink that can always accept input.
Both systems rely on stream bindings being used linearly to ensure correctness, including boundedness of buffers and ordering of effects.
Operations over sources are expressed fairly naturally compared to streams, for example the @zip@ combinator has the type @Source a -> Source b -> Source (a,b)@.
Sinks, however, are co-variant, and operations must be performed somewhat backwards, so that the @unzip@ combinator takes the two output sinks to push into and returns a new sink that pushes into these.
It has the type @Sink a -> Sink b -> Sink (a,b)@.
This system requires the streaming computation to be manually split into sources and sinks, and be joined together by a loop that `drains' values from the source and pushes them into the sink.

The duality between pull and push arrays has also been explored in Obsidian by \citep{claessen2012expressive} and later in \citep{svensson2014defunctionalizing}.
Here the distinction is made for the purpose of code generation for GPUs rather than fusion, as operations such as appending pull arrays require conditionals inside the loop, whereas using push arrays moves these conditionals outside the loop.

Streaming IO libraries have blossomed in the Haskell ecosystem, generally based on Iteratees \cite{kiselyov2012iteratees}.
Libraries such as @conduit@ \cite{hackage:conduit}, @enumerator@ \cite{hackage:enumerator}, @machines@ \cite{hackage:machines} and @pipes@ \cite{hackage:pipes} are all designed to write stream computations with bounded buffers.
However, these libraries provide no fusion guarantees, and as such programs tend to be written over chunks of data to make up for the communication overhead.
For the most part they support only straight-line computations, with only limited forms of branching.

In relation to process calculi, synchronised product has been suggested as a method for fusing Kahn process networks together~\cite{fradet2004network}, but does not appear to have been implemented or evaluated.
The synchronised product of two processes allows either process to take independent or local steps at any time, but shared actions, such as when both processes communicate on the same channel, must be taken in both processes at the same time.
This is a much simpler fusion method than ours, but is also much stricter.
When two processes share multiple channels, synchronised product will fail unless both processes read the channels in exactly the same order.
Our system can be seen as an extension of synchronised product that allows some leeway in when processes must take shared steps: they do not have to take shared steps at the same time, but if one process lags behind the other, it must catch up before the other one gets too far ahead.

Synchronous languages such as LUSTRE~\cite{halbwachs1991synchronous}, Lucy-n~\cite{mandel2010lucy} and SIGNAL~\cite{le2003polychrony} all use some form of clock calculus and causality analysis to ensure that programs can be statically scheduled with bounded buffers.
These languages describe \emph{passive} processes where values are fed in to streams from outside environments, such as data coming from sensors.
In this case, the passive process has no control over the rate of input coming in, and if they support multiple input streams, they must accept values from them in any order.
In contrast, the processes we describe are \emph{active} processes that have control over the input that is coming in.
This is necessary for combinators such as mergesort-style @merge@, as well as @append@.
Note that in the synchronous language literature, it is common to refer to a different merge operation, also known as @default@, which computes a stream that is defined whenever either input is defined.

Synchronous dataflow (not to be confused with synchronous languages above) is a dataflow graph model of computation where each dataflow actor has constant, statically known input and output rates.
The main advantage of synchronous dataflow is that it is simple enough for static scheduling to be decidable, but this comes at a cost of expressivity.
StreamIt~\cite{thies2002streamit} uses synchronous dataflow for scheduling when possible, otherwise falling back to dynamic scheduling~\cite{soule2013dynamic}.
Boolean dataflow and integer dataflow~\cite{buck1993scheduling,buck1994static} extend synchronous dataflow with boolean and integer valued control ports, and attempt to recover the structure of ifs and loops from select and switch actors.
These systems allow some dynamic structures to be scheduled statically, but are very rigid and only support limited control flow structures: it is unclear how merge or append could be scheduled by this system.
Finite state machine-based scenario aware dataflow (FSM-SADF)~\cite{stuijk2011scenario,van2015scenario} is still quite expressive compared to boolean and integer dataflow, while still ensuring static scheduling.
A finite state machine is constructed, where each node of the FSM denotes its own synchronous dataflow graph.
The FSM transitions from one dataflow graph to another based on control outputs of the currently executing dataflow graph.
For example, a filter is represented with two nodes in the FSM.
The dataflow graph for the initial state executes the predicate, and the value of the predicate is used to determine which transition the FSM takes: either the predicate is false and the FSM stays where it is, or the predicate is true and moves to the next state.
The dataflow graph for the next state emits the value, and moves back to the first state.
This does appear to be able to express value-dependent operations such as merge, but lacks the composability - and familiarity - of combinators.

% StreamIt:
% Only allows limited splits and joins: round robin and duplication for splits, round robin and combination for joins. 
% Does not support fully general graphs - instead using combinators to introduce a (split/join) and a combinator for a feedback loop.
% 
% Parameterized dataflow (PDF),  \cite{bhattacharya2001parameterized}
% Schedulable parametric dataflow (SPDF),  \cite{fradet2012spdf}