### 2017-02-25 2017mergingmerges/section/XX-Evaluation.tex
End S2 where it is, that is enough intuition before going onto the defs

%!TEX root = ../Main.tex

\section{Evaluation}
\label{s:Evaluation}

\TODO{Discuss the performance of the system. Why this is a good idea. Give graphs that show the number of states is sensible when more processes are fused in.}

\subsection{Optimisation}
After the algorithm has completed fusing both processes it may be possible to eliminate redundant @jump@, depending on how the affected variables are used by other instructions.

\TODO{Discuss combining jump instructions here. Use this to motivate the need for drop instructions.}