### 2023-05-11 2023pipit/figures/core-grammar.tex

\begin{figure}
  \[
  \begin{array}{lrlll}
    e, e' & := & v ~|~ x ~|~ e~e' \\
          & | & \xfby{v}{e} & | & \xthen{e}{e'} \\
          & | & \xrec{x}{e[x]} & | & \xcheck{e} \\
          & | & \multicolumn{3}{l}{\xlet{x}{e}{e'[x]}} \\
    \\
    v & := & \multicolumn{3}{l}{n \in \mathbb{N} ~|~ b \in \mathbb{B} ~|~ r \in \mathbb{R} ~|~ \hdots ~|~ \mlamX} \\
    % \sigma & := & \sgl{\ov{x \mapsto v}} \\
    % \Sigma & := & \sigma ~|~ \Sigma;\sigma \\
    %   \\
    % \tau, \tau' & := & \multicolumn{3}{l}{\mathbb{N} ~|~ \mathbb{B} ~|~ \hdots ~|~ \tau \to \tau'} \\
    % \Gamma & := & \cdot ~|~ x : \tau, \Gamma \\
    \end{array}
  \]
  \caption{Pipit language grammar including expressions~($e$); and values and host functions embedded from the \fstar{} meta-language~($v$).}
  \label{f:core-grammar}
\end{figure}