### 2023-12-20 2024pipit/section/35-Primitive-Tables.tex

\subsection{Primitive tables}
\label{s:core:primitive-tables}

The core language presented in \autoref{f:core-grammar} has a restricted set of primitives.
Our actual implementation is parameterised by the set of primitive types and operations.

In addition to a standard deeply-embedded primitive table which supports integers, tuples, booleans and reals, we also provide a shallowly-embedded primitive table, where operations can be arbitrary pure \fstar{} functions.
This shallow embedding also allows straightforward integration of \fstar{} types and functions, as well as proofs inside streaming functions.

This parameterisation serves mainly as a convenience.
For the formalisation presented here, we assume a simple static set of primitives.

In our current implementation, we cannot encode primitives that have preconditions, as the preconditions are checked eagerly by \fstar{}, before anything is known about the surrounding streaming values.

To work around this limitation, we encode certain primitive preconditions as implications on the postcondition.
For example, to encode the array indexing operation in \autoref{s:motivation}, we use an unsafe variant of the indexing function:
\begin{tabbing}
  @MM@\= @MM@ \= \kill
  @val@ Array.index_unsafe ($\textit{array}: \mbox{const_array}~\alpha$) ($\textit{index}$: $\NN$): \\
  \> $v: \alpha \{ \textit{index} < \mbox{Array.length}~\textit{array} \implies v = \mbox{Array.index}~\textit{array}~\textit{index} \}$
\end{tabbing}

By wrapping this unsafe index function inside a contract that requires the index to be valid, we can prove the postcondition on the streaming level.

In the future, we plan to investigate allowing preconditions on non-streaming functions, so that safe variants can be used directly.