### 2024-01-10 2024pipit/section/35-Primitive-Tables.tex

% KILL I don't think this is really useful

% \subsection{Primitive tables}
% \label{s:core:primitive-tables}

% The core language presented in \autoref{f:core-grammar} has a restricted set of primitives.
% Our actual implementation is parameterised by the set of primitive types and operations.

% In addition to a standard deeply-embedded primitive table which supports integers, tuples, booleans and reals, we also provide a shallowly-embedded primitive table, where operations can be arbitrary pure \fstar{} functions.
% This shallow embedding also allows straightforward integration of \fstar{} types and functions, as well as proofs inside streaming functions.

% This parameterisation serves mainly as a convenience.
% For the formalisation presented here, we assume a simple static set of primitives.