### 2024-01-18 2024pipit/Main.tex
update evaluation

% \documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
% To anonymise:
\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate,anonymous]{lipics-v2021}

% \pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

\bibliographystyle{plainurl}

\ifx\authoranonymous\relax
\newcommand{\GITHUBURL}{(anonymised)}
\newcommand{\COFFEEURL}{(anonymised)}
\else
\newcommand{\GITHUBURL}{\url{https://github.com/songlarknet/pipit}}
\newcommand{\COFFEEURL}{\url{https://youtu.be/6IybbQFPOl8}}
\fi

% is this package necessary?
\usepackage{mathtools}
\usepackage{mathpartir}

\usepackage{stmaryrd}

% checked box
\usepackage{wasysym}
\usepackage{underscore}

\usepackage{style/utils}
\usepackage{style/code}
\usepackage{style/proof}
\usepackage{style/keywords}
\usepackage{style/judgements}

\title{Pipit on the Post: Proving Pre- and Post-conditions of Reactive Systems}

\author{Amos Robinson}{Australian National University, Canberra, Australia}{amos.robinson@anu.edu.au}{https://orcid.org/0009-0004-4837-4981}{}

\author{Alex Potanin}{Australian National University, Canberra, Australia}{alex.potanin@anu.edu.au}{https://orcid.org/0000-0002-4242-2725}{}

\authorrunning{A. Robinson and A. Potanin}
\Copyright{Amos Robinson and Alex Potanin}
% \ccsdesc[300]{Computer systems organization~Embedded software}
\ccsdesc[500]{Computer systems organization~Real-time languages}
\ccsdesc[500]{Theory of computation~Program verification}
% \ccsdesc[100]{Theory of computation~Modal and temporal logics}
\ccsdesc[500]{Software and its engineering~Specialized application languages}
\keywords{Lustre, streaming, reactive, verification}

% \category{}

\begin{document}

\maketitle

\begin{abstract}
  Reactive languages such as Lustre and Scade are used to implement safety-critical control systems; proving such programs correct and having the proved properties apply to the compiled code is therefore equally critical.
  We introduce Pipit, a small reactive language embedded in \fstar{}, designed for verifying control systems and executing them in real-time.
  Pipit includes a verified translation to transition systems; by reusing \fstar{}'s existing proof automation, certain safety properties can be automatically proved by k-induction on the transition system.
  Pipit can also generate executable code in a subset of \fstar{} which is suitable for compilation and real-time execution on embedded devices.
  The executable code is deterministic and total and preserves the semantics of the original program.
\end{abstract}

% enable @verbatim@ syntax
\makeatactive

\input{section/01-Introduction.tex}
\input{section/20-Motivation.tex}
\input{section/30-Core.tex}
\input{section/35-Primitive-Tables.tex}
\input{section/40-Transition.tex}
\input{section/50-Extraction.tex}
\input{section/60-Evaluation.tex}
\input{section/70-Related.tex}
\input{section/80-Conclusion.tex}

\bibliography{Main}

% \appendix

\end{document}
\endinput